Teaching Dossier

### 2019-11-21 CUT/teaching-philosophy.tex
wrote up on first teaching observation

On a question-by-question level it can be difficult to coax responses out of students. It is important that I ensure I wait sufficiently long for students to respond. I use a ten second count in my head, or take a sip of water, to make sure I'm pausing long enough for a student to both work out the answer and work up the confidence to respond. Most of the time no hands go up immediately, but after a waiting period some hands will go up. Although sometimes students are not responding because they genuinely do not know the answer. When I get a sense this is happening, I will break the question down for them into smaller and easier parts. At some level I will find something simple enough that someone in the class will know the answer and be willing to respond. I prefer to not simply give the answer and move on. Although it may feel like the best thing to do to preserve momentum and save time, the fact that students are not responding means we have identified a gap in their understanding and it is important to drill down on this gap and help students bridge it piece by piece.
\subsection{Outside Lecture}

\subsection{Assessment}
\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}
%highlights from 1st observation
%highlights from 2nd observation with a narrative about improving

@@

\appendix
\section*{Appendices}
%Full evaluations
%Observation Reports
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{}

>>>>

On a question-by-question level it can be difficult to coax responses out of students. It is important that I ensure I wait sufficiently long for students to respond. I use a ten second count in my head, or take a sip of water, to make sure I'm pausing long enough for a student to both work out the answer and work up the confidence to respond. Most of the time no hands go up immediately, but after a waiting period some hands will go up. Although sometimes students are not responding because they genuinely do not know the answer. When I get a sense this is happening, I will break the question down for them into smaller and easier parts. At some level I will find something simple enough that someone in the class will know the answer and be willing to respond. I prefer to not simply give the answer and move on. Although it may feel like the best thing to do to preserve momentum and save time, the fact that students are not responding means we have identified a gap in their understanding and it is important to drill down on this gap and help students bridge it piece by piece.

\subsection{Outside Lecture}

\subsection{Assessment}

\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}
One element of the CUT teaching program I took alongside my PhD was having lectures I give upervised by an expert in teaching working with the university's Centre for Teaching Excellence (CTE). I completed two teaching observations through the program. Below I summarize the feedback I received from these observations.
%highlights from 1st observation
\subsubsection*{First Teaching Observation}
\paragraph{}
My first observation was on the guest lecture I gave for Introduction to Game Theory in Fall $2018$ on Correlated Equilibria. The observation was conducted by Shahrukh Athar, a Graduate Instructional Developer working for the CTE. I have included the relevant sections of the full report in Appendix \ref{sec:observation-report-1}. The feedback in the report was divided in \emph{aspects to maintain}, where strengths of teaching were identified, and \emph{targets for change}, where opportunities for improvement were identified.
\paragraph{Aspects to Maintain} The first strength Shahrukh identified was my use of questions and comprehension checks. He appreciated the variety of checks I used, as well as the split between content-specific checks and general purpose open-ended questions. He commented that ``\emph{All of this is
fantastic as it demonstrates a learner-centric approach to teaching rather than a teacher-centric
one. It shows that you want to take your learners along with you as the lesson progresses and
are willing to pause and clarify confusions before moving on.}''

Shahrukh also found my lesson structure and transitions to be very strong, saying he will ``\emph{encourage [me] to continue
following such a structured approach in the delivery of [my] future lessons}''. He appreciated my use of an outline at the start of the lesson the students could follow along with, and found my method of using comprehension checks to indicate upcoming transitions to the next component of a lesson to be a highly effective technique.

Lastly, Shahrukh really enjoyed the high degree of interactivity in my lessons. Citing both the instructor-student questions throughout the lesson, and student-student interaction which occurred during the post-test. He particularly liked the latter, ``\emph{ I specifically want to celebrate here is the learner-learner
activity that you had towards the end of the lesson}.'' The activity he refers to is a computation of a correlated equilibrium that the students were asked to perform in pairs during the last $10$ minutes of the lecture.

\paragraph{Targets for Change} Shahrukh identified for me a couple of areas in which I could improve my teaching. The first area he identified involved some practical matters he dubbed ``awareness of learner needs''. He sat at the back of the classroom for the lecture, and found that from his position my handwriting was ``\emph{a touch small}''. The fix for this is easy, and I have since adapted it in all my lessons. Before the lecture starts I write some sentences and pieces of mathematics on the board, then head to the back of the room to verify that the font size is legible from that far away. Shahrukh also observed that there were a couple of noise disruptions caused by students walking in the halls outside the classroom since I left the doors open. I had initially done this to be welcoming to students who arrive late, but have since starting closing doors at the start of the lecture after verifying there are no students rushing through the hall trying to reach the classroom.

Shahrukh noticed a habit I had during the lecture of speaking while writing and facing the board. While there were no issues with my voice projecting, it created a visual disconnect with the students. I have since adopted his strategy of ``Write-Turn-Talk'', where I am silent during writing on the board and turn around to explain what I just wrote to my students. A major strength of this technique is that it does not ask students to process two streams of information at once (my voice and my writing), but rather they can concentrate on copying what I a writing as I write it and then concentrate on the verbal explanation I give after it is written down. In my second teaching observation I specifically asked the observer to verify that I had successfully implemented this technique.

\subsubsection*{Second Teaching Observation}
\paragraph{}My second observation was conducted by Caelan Wang, a Graduate Educational Developer with the CTE. The lecture to be observed was one my regularly scheduled lectures as the instructor for Deterministic OR Models in Spring $2019$, covering cutting plane methods for solving integer programs. Caelan was especially qualified to observe this lecture since she was a graduate from the same department I am pursuing my PhD in, and knew the topic well.

\paragraph{Aspects to Maintain}

\paragraph{Targets for Change}
%highlights from 2nd observation with a narrative about improving

@@

\appendix
\section*{Appendices}
%Full evaluations
%Observation Reports
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{Teaching Observation Report 1}\label{sec:observation-report-1}
